---
title: "SDS358_Project_EDA"
author: "Brent Bouslog"
date: "10/15/2020"
output: pdf_document
---

```{r setup, include=FALSE}
library(tidyverse)
knitr::opts_chunk$set(echo = TRUE)
require("knitr")
opts_knit$set(root.dir = "~/RStudio")
```

## Visualizing the Data Table

```{r show_orig, layout="l-body-outset"}
mydata <- read.csv("nest_predation_dataset.csv")
kable(head(mydata))
```

## Visualizing Clean Data Table

```{r show, layout="l-body-outset"}
mydata_clean <- mydata[c("Lat","Long","IncubationPeriod.days.", "NestlingPeriod.days.", "ClutchSize", "demelevation", "temptrop2", "PercentageSuccessfulNests")]
mydata_clean <- drop_na(mydata_clean)
kable(head(mydata_clean))
```

## Univariate Statistics

```{r stats}
summary.extended <- mydata_clean %>%
  select(Lat,Long,IncubationPeriod.days.,NestlingPeriod.days.,ClutchSize,demelevation,PercentageSuccessfulNests) %>%
  psych::describe(fast = TRUE) %>%
  as_tibble(rownames="rowname") %>%
  print(summary.extended)
```

## Visualization of Latitue

```{r lat}
ggplot(mydata_clean, aes(Lat)) +
  geom_histogram(fill="lightblue", color="blue", aes(y = stat(count) / sum(count))) +
  labs(title ="Histrogram of Latitude Observations", x = "Latitude (in degrees)", y = "Relative frequency") +
  theme_classic()
```

From this graph we can tell that the distribution of latitudes is bi-modal and symmetric with many observations falling around 0 degrees and 40 degrees. This puts many of our observations very close to the equator.

## Visualization of Longitude

```{r long}
ggplot(mydata_clean, aes(Long)) +
  geom_histogram(fill="lightblue", color="blue", aes(y = stat(count) / sum(count))) +
  labs(title ="Histrogram of Longitude Observations", x = "Longitude (in degrees)", y = "Relative frequency") +
  theme_classic()
```

From this graph we can tell that the distribution of longitudes is unimodal and symmetric with most observations falling around -80 degrees. Due to the fact that none of our observations had longitudes greater than 0 we can say that all of our observations are in the western hemisphere.

## Visualization of Latitude vs. Longitude

```{r lat_log}
ggplot(mydata_clean, aes(x=Long, y=Lat)) + 
  geom_point() +
  labs(title ="Latitude vs. Longitude", x = "Longitude (in degrees)", y = "Latitude (in degrees)") +
  theme_classic()
```

This graph emulates a map of the locations of all the nests. From this graph we can tell that most nests are from North, Central, and South America.

## Visualization of Incubation Period (days)

```{r incubation}
ggplot(mydata_clean, aes(IncubationPeriod.days.)) +
  geom_histogram(fill="lightblue", color="blue", aes(y = stat(count) / sum(count))) +
  labs(title ="Histrogram of Incubation Lengths", x = "Incubation Length (in days)", y = "Relative frequency") +
  theme_classic()
```

From this graph we can tell that the distribution of incubation lengths is unimodal and skewed right with many observations falling between 10 and 15 days with a few observations of incubation lengths greater than 25 days.

## Visualization of Nestling Period (days)

```{r nestling}
ggplot(mydata_clean, aes(NestlingPeriod.days.)) +
  geom_histogram(fill="lightblue", color="blue", aes(y = stat(count) / sum(count))) +
  labs(title ="Histrogram of Nestling Lengths", x = "Nestling Length (in days)", y = "Relative frequency") +
  theme_classic()
```

From this graph we can tell that the distribution of Nestling Lengths is unimodal and skewed right with many observations falling around 10 days with a few observations greater than 30 days.

## Visualization of Clutch Size

```{r clutch}
ggplot(mydata_clean, aes(ClutchSize)) +
  geom_histogram(fill="lightblue", color="blue", aes(y = stat(count) / sum(count))) +
  labs(title ="Histrogram of Clutch Sizes", x = "Clutch Size (in number of eggs)", y = "Relative frequency") +
  theme_classic()
```

From this graph we can tell that clutch size is unimodal and skewed right with most clutches having close to 2 eggs with a few having as many as 6 eggs.

## Visualization of Altitude (m)

```{r alt}
ggplot(mydata_clean, aes(demelevation)) +
  geom_histogram(fill="lightblue", color="blue", aes(y = stat(count) / sum(count))) +
  labs(title ="Histrogram of Nest Altitude", x = "Nest Altitude (in m above sea level)", y = "Relative frequency") +
  theme_classic()
```

From this graph we can tell that the distribution of nest altitudes is unimodal and skewed right with many altitudes being very close to sea level with only a few altitudes in the mountains around 2000-3000 m about sea level.

## Visualization of Habitat Type

```{r temp}
ggplot(mydata_clean, aes(temptrop2)) + 
  geom_bar(aes(y = stat(count) / sum(count))) +
  labs(title ="Bar Graph of Habitat Types", x = "Habitat Type", y = "Relative frequency") +
  theme_classic()
```

From this graph we can tell that the majority of the nests come from northern temperate locations, followed closely behind by tropical locations, with only a few nests coming from southern temperate locations.

## Visualization of Nest Success Rate

```{r success}
ggplot(mydata_clean, aes(PercentageSuccessfulNests)) +
  geom_histogram(fill="lightblue", color="blue", aes(y = stat(count) / sum(count))) +
  labs(title ="Histrogram of Percentage of Successful Nests", x = "Percentage of Successful Nests", y = "Relative frequency") +
  theme_classic()
```

From this graph we can tell that the success rate of nests is unimodal and symmetric with many success rates falling around the 50% mark with only a few close to 0% and 100%.